High-Efficiency, Small-Footprint Couplers Between Arbitrary Nanophotonic Waveguide Modes in Two Dimensions

### 2011-12-15 presentations/ob1_wg_paper/paper.tex
Made changes to ob1 wg paper.

\begin{document}
\title{Objective-First Design of Nanophotonic Waveguide Couplers}
\author{Jesse Lu and Jelena Vu\v{c}kovi\'{c}}
\address{Stanford University, Stanford, California, USA.}
\email{jesselu@stanford.edu}

\maketitle
\begin{abstract}
We introduce an ``objective-first'' approach to the design of
    nanophotonic waveguide couplers.
Our method is
    computationally fast (15 minutes on a single-core personal computer), 
    requires no trial-and-error,
    does not require guessing a good starting design,
    can be applied to arbitrary waveguide modes,
    and generates devices with high coupling efficiencies (typically 95\%)
    and small footprints (1-4 square vacuum wavelengths).
\end{abstract}
\ocis{230.7370, 130.3990.}

@@

% Why is waveguide mode conversion important?
\subsection{The importance of waveguide mode conversion}
Optical mode conversion, 
    the efficient transfer of photons from one guided mode to another,
    is a fundamental requirement in nanophotonics.
Efficient conversion between waveguides modes
    is critical in the many cases, including:
    \begin{enumerate}
        \item Coupling to and from optical fiber\cite{fibergrating}, 
        to communicate with the outside world.
    \item Coupling between various nanophotonic waveguides, 
        since different waveguides are best suited for different applications.
        For example, ridge waveguides seem ideal for 
            low-loss transport\cite{ridge},
            but other waveguides, 
            such as photonic crystal waveguides or slot waveguides,
            may be better suited for slow-light\cite{pcslow} 
            or nonlinear optical devices based on 
            localized field intensities\cite{slotfocus}.
    \item Coupling between different materials systems such as
       passive, active\cite{active}, and metallic\cite{metallic} devices. 
    \end{enumerate}
Although efficient waveguide coupling is essential in any nanophotonic system,
    practical methods to design couplers have not yet been developed.

@@

\section{Results}
We apply our method to five practically interesting problems 
    that are hard to solve using other available methods: 
    \begin{enumerate}
    \item a coupler between waveguides of different refractive index and width,
    \item a coupler between waveguide modes of different order and symmetry,
    \item a coupler between waveguides that confine light 
        using different principles 
        (index guided vs. distributed Bragg reflection guided), 
        i.e., between a slab waveguide and a photonic crystal fiber. 
    \item a coupler from a dielectric to a plasmonic metal-insulator-metal 
        waveguide, and
    \item a coupler from a dielectric waveguide to a (plasmonic) metal wire.
    \end{enumerate}

The corresponding designs as well as the boundary values which make up the
    respective design objectives ($H_z^\text{perfect}$) 
    are shown in Figs.~\ref{fig:fiber}, \ref{fig:mode} and \ref{fig:aircore}.
As initial designs, we uniformly choose $\epsilon = 9$ throughout 
    the whole computational region.
These results demonstrate that our method
    \begin{itemize}
    \item is computationally fast, 
        requiring only 15 minutes on a single-core personal computer;
    \item generates compact devices with footprints of only
        1-4 square vacuum wavelengths;
    \item generates highly efficient devices with  
        typical coupling efficiencies of $95\%$;
    \item can be applied to arbitrary input and output modes,
        as seen from the diverse selection of desired output waveguide modes;
    \item does not require a good initial design.
        The initial design was simply $\epsilon = 9$ everywhere
            (a somewhat arbitrary guess, other values work as well),
            which in the case of Figs.~\ref{fig:mode} and \ref{fig:aircore}
            produces an initially non-functional device 
            (coupling efficiency of $0\%$).

    \end{itemize}

\begin{figure}[htbp]
    \centering
    \includegraphics[width=\textwidth]{1} 
    \caption{Coupler from a narrow, high-index ($\epsilon=12.25$) waveguide to
            a wide, low-index ($\epsilon=2.25$) waveguide. 
        The $H_z^\text{perfect}$ boundary values used as 
            the design objective are plotted in the upper-left quadrant, 
            the generated structure is shown in the upper-right quadrant, and
            the simulated $H_z$ fields of the device are shown 
            in the bottom two plots.
        The computed efficiency of the coupler is high, $98.2\%$, and
            the device is also extremely compact, 
            convering only $36 \times 66$ grid points,
            where the vacuum wavelength is 42 grid points.
        Computation time was 15 minutes on a personal computer.}
    \label{fig:fiber}
\end{figure}
\begin{figure}[htbp]
    \centering
    \includegraphics[width=\textwidth]{2} 
    \caption{Coupler that converts the fundamental waveguide mode to the
            second-order waveguide mode
            (a movie of the design progress is included as a
            supplementary material).
        This problem is quite difficult since the two modes are of 
            opposite symmetry.
        For example, adiabatic approaches cannot be applied to this case.
        However, our method produces a device 
            (which has the same dimensions and vacuum wavelength as 
            Fig.~\ref{fig:fiber})  
            with a coupling efficiency of $97.0\%$. 
        Computation time was 15 minutes on a personal computer.
        }
    \label{fig:mode}
\end{figure}
\begin{figure}[htbp]
    \centering
    \includegraphics[width=\textwidth]{3}
    \caption{Coupler between a dielectric slab waveguide to 
            an air-core waveguide.
        Here, not only are the modes of opposite symmetry,
            but the output waveguide operates on a fundamentally different
            principle (distributed reflection) than the input waveguide 
            (index guided).
        The device still achieves an efficiency of $85.1\%$, demonstrating the
            versatility of our method.
        The vacuum wavelength is 25 grid points, 
            while the device footprint is still $36 \times 66$ grid points.
        Computation time was 15 minutes on a personal computer.
        }
        \label{fig:aircore}
\end{figure}
\begin{figure}[htbp]
    \centering
    \includegraphics[width=\textwidth]{4}
    \caption{Coupler between a dielectric slab waveguide to 
            a plasmonic metal-insulator-metal waveguide.
        The efficiency of the device is $92.9\%$ and 
            has the same wavelength and footprint as the device in
            Fig.~\ref{fig:aircore}.
        Computation time was 15 minutes on a personal computer.
        }
        \label{fig:mim}
\end{figure}
\begin{figure}[htbp]
    \centering
    \includegraphics[width=\textwidth]{5}
    \caption{Coupler between a dielectric slab waveguide to 
            a plasmonic wire waveguide.
        The efficiency of the device is $93.2\%$ and 
            has the same wavelength and footprint as the device in
            Fig.~\ref{fig:aircore}.
        Computation time was 15 minutes on a personal computer.
        }
        \label{fig:wire}
\end{figure}

@@

We then apply an objective-first approach to the design of 
    five practical nanophotonic waveguide couplers which are difficult, 
    at best, to solve with existing methods. 
We show that our method produces
    high-efficiency designs ($\sim 95\%$ efficiency) 
    in small footprints ($\sim 1$ square vacuum wavelength),
    is computationally fast (15 minutes on a single-core personal computer), and
    does not require trial-and-error, or even 
    a good starting design.\\

>>>>

\begin{document}
\title{High-Efficiency, Small-Footprint Couplers Between 
    Arbitrary Nanophotonic Waveguide Modes in Two Dimensions}
\author{Jesse Lu and Jelena Vu\v{c}kovi\'{c}}
\address{Stanford University, Stanford, California, USA.}
\email{jesselu@stanford.edu}

\maketitle
\begin{abstract}
We develop a design algorithm which can design 
    high efficiency ($\sim$95\%), 
    small-footprint (1-4 square vacuum wavelengths)
    couplers between arbitrary nanophotonic waveguide modes.
Such a capability is extremely useful since it fundamentally allows for the
    design of all linear optical devices.
Our ``objective-first'' method is
    computationally fast (15 minutes on a single-core personal computer), 
    requires no trial-and-error, and
    does not require guessing a good starting design.
\end{abstract}
\ocis{230.7370, 130.3990.}

@@

% Why is waveguide mode conversion important?
\subsection{The importance of waveguide mode conversion}
Optical mode conversion, 
    the efficient transfer of photons from one guided mode to another,
    is a fundamental requirement in nanophotonics.
In fact, the problem of converting between nanophotonic modes is essentially
    the function of \emph{all} linear nanophotonic devices.
For instance, efficient conversion between waveguides modes
    is useful for:
    \begin{enumerate}
        \item Coupling to and from optical fiber\cite{fibergrating}, 
        to communicate with the outside world.
    \item Coupling between various nanophotonic waveguides, 
        since different waveguides are best suited for different applications.
        For example, ridge waveguides seem ideal for 
            low-loss transport\cite{ridge},
            but other waveguides, 
            such as photonic crystal waveguides or slot waveguides,
            may be better suited for slow-light\cite{pcslow} 
            or nonlinear optical devices based on 
            localized field intensities\cite{slotfocus}.
    \item Coupling between different materials systems such as
       passive, active\cite{active}, and metallic\cite{metallic} devices. 
    \end{enumerate}
Although efficient waveguide coupling is essential in any nanophotonic system,
    practical methods to design couplers have not yet been developed.

@@

\section{Results}
We now present evidence that the method outlined above 
    is indeed able to design coupling structures between essentially
    any two nanophotonic waveguide modes.
To do so, we apply our method to the following five problems:
    \begin{enumerate}
    \item coupling between waveguides of different refractive index and width,
    \item coupling between waveguide modes of different order and symmetry,
    \item coupling between waveguides that confine light 
        using different principles 
        (index guided vs. distributed Bragg reflection guided), 
        i.e., between a slab waveguide and a photonic crystal fiber. 
    \item coupling from a dielectric to a plasmonic metal-insulator-metal 
        waveguide, and
    \item coupling from a dielectric waveguide to a (plasmonic) metal wire,
    \end{enumerate}
    the results of which are shown in Figs.~\ref{fig:fiber}-\ref{fig:wire}.
\begin{figure}[htbp]
    \centering
    \includegraphics[width=0.8\textwidth]{1} 
    \caption{Coupler from a narrow, high-index ($\epsilon=12.25$) waveguide to
            a wide, low-index ($\epsilon=2.25$) waveguide. 
        The $H_z^\text{perfect}$ boundary values used as 
            the design objective are plotted in the upper-left quadrant, 
            the generated structure is shown in the upper-right quadrant, and
            the simulated $H_z$ fields of the device are shown 
            in the bottom two plots.
        The computed efficiency of the coupler is high, $98.2\%$, and
            the device is also extremely compact, 
            convering only $36 \times 66$ grid points,
            where the vacuum wavelength is 42 grid points.
        Computation time was 15 minutes on a personal computer.}
    \label{fig:fiber}
\end{figure}
\begin{figure}[htbp]
    \centering
    \includegraphics[width=0.8\textwidth]{2} 
    \caption{Coupler that converts the fundamental waveguide mode to the
            second-order waveguide mode
            (a movie of the design progress is included as a
            supplementary material).
        This problem is quite difficult since the two modes are of 
            opposite symmetry.
        For example, adiabatic approaches cannot be applied to this case.
        However, our method produces a device 
            (which has the same dimensions and vacuum wavelength as 
            Fig.~\ref{fig:fiber})  
            with a coupling efficiency of $97.0\%$. 
        Computation time was 15 minutes on a personal computer.
        }
    \label{fig:mode}
\end{figure}
\begin{figure}[htbp]
    \centering
    \includegraphics[width=0.8\textwidth]{3}
    \caption{Coupler between a dielectric slab waveguide to 
            an air-core waveguide.
        Here, not only are the modes of opposite symmetry,
            but the output waveguide operates on a fundamentally different
            principle (distributed reflection) than the input waveguide 
            (index guided).
        The device still achieves an efficiency of $85.1\%$, demonstrating the
            versatility of our method.
        The vacuum wavelength is 25 grid points, 
            while the device footprint is still $36 \times 66$ grid points.
        Computation time was 15 minutes on a personal computer.
        }
        \label{fig:aircore}
\end{figure}
\begin{figure}[htbp]
    \centering
    \includegraphics[width=0.8\textwidth]{4}
    \caption{Coupler between a dielectric slab waveguide to 
            a plasmonic metal-insulator-metal waveguide.
        The efficiency of the device is $92.9\%$ and 
            has the same wavelength and footprint as the device in
            Fig.~\ref{fig:aircore}.
        Computation time was 15 minutes on a personal computer.
        }
        \label{fig:mim}
\end{figure}
\begin{figure}[htbp]
    \centering
    \includegraphics[width=0.8\textwidth]{5}
    \caption{Coupler between a dielectric slab waveguide to 
            a plasmonic wire waveguide.
        The efficiency of the device is $93.2\%$ and 
            has the same wavelength and footprint as the device in
            Fig.~\ref{fig:aircore}.
        Computation time was 15 minutes on a personal computer.
        }
        \label{fig:wire}
\end{figure}

These results clearly demonstrate that our method
    \begin{itemize}
    \item can be applied to arbitrary input and output modes,
        as seen from the diverse selection of desired output waveguide modes;
    \item generates highly efficient devices with  
        typical coupling efficiencies of $95\%$;
    \item generates compact devices with footprints of only
        1-4 square vacuum wavelengths;
    \item does not require a good initial design.
        The initial design was simply $\epsilon = 9$ everywhere
            (a somewhat arbitrary guess, other values work as well),
            which in almost every case
            produces an initially non-functional device 
            (coupling efficiency of $0\%$);
    \item is computationally fast, 
        requiring only 15 minutes on a single-core personal computer.
    \end{itemize}

Critically, these results form a key advancement in our understanding of 
    nanophotonic design; namely, 
    that it is indeed possible to efficiently design
    virtually any linear optical device using an objective-first approach.

@@

We then apply an objective-first approach to the design of 
    five practical nanophotonic waveguide couplers which are difficult, 
    at best, to solve with existing methods. 
We show that our method produces
    high-efficiency designs ($\sim 95\%$ efficiency) 
    in small footprints ($\sim 1$ square vacuum wavelength)
    between arbitrary waveguide modes.
Furthermore, out method
    is computationally fast (15 minutes on a single-core personal computer), and
    does not require trial-and-error, or even 
    a good starting design.\\
